State-space PTA

### 2023-03-13 docs/manuscript/paper.tex
paper corrections to fundamental equations

\section{Model}
The intrinsic pulsar frequency $f$ evolves as (See Vargas and Melatos)
\begin{equation}
	\frac{df}{dt} = -\gamma	 [f - f_{\rm EM} (t)] + \dot{f}_{\rm EM} + \xi(t)
	\label{eq:frequency_evolution}
\end{equation}
where $f_{\rm EM}$ is the solution of the electromagnetic spindown equation, $\gamma$ a proporionality constant, and $\xi(t)$ a white noise process that satisfies,
\begin{equation}
	\langle \xi(t) \xi(t') \rangle = \sigma^2 \delta(t - t')
\end{equation}
Over the timescales that we are interested in, we can express the EM spindown simply as
\begin{equation}
	f_{\rm EM} (t) = f_{\rm EM}(0) + \dot{f}_{\rm EM} (0)t
\end{equation}  
Completely, the frequency evolution is then
\begin{equation}
	\frac{df}{dt} = -\gamma	 [f - f_{\rm EM}(0) - \dot{f}_{\rm EM} (0)t] + \dot{f}_{\rm EM} (0) + \xi(t)
	\label{eq:frequency_evolution_expanded}
\end{equation}
This intrinsic frequency can be related to a measured frequency as 
\begin{equation}
	f_M = f g(\bar{\theta},t) + N_M
\end{equation}
where $g(\theta,t)$ ("measurement function")is a function of some parameters, $\bar{\theta}$ and time $t$, whilst $N_M$ is a Gaussian measurement noise that satisfies 
\begin{equation}
	\langle N_M(t) N_M(t') \rangle = \Sigma^2 \delta(t - t')
\end{equation}
Explicitly, it can be shown that the measurement function is,
\begin{equation} \label{eq:final}
	g(\bar{\theta},t) = 1 - \frac{1}{2} \frac{h_{ij}(t) q(t)^i q(t)^j}{1 + \bar{n} \cdot \bar{q}(t)}[1 - e^{i \Omega(1+\bar{n} \cdot \bar{q}(t)) d}]
\end{equation}
There are a few terms to unpack and define in this equation. $q(t)$ is the vector that connects the Earth and the pulsar, $n$ the vector that connects the Earth and the GW source, $\Omega$ the (constant) angular frequency of the GW and $d$ the Earth-pulsar distance. $h_{ij}(t)$ is the metric perturbation due to the GW $= g_{ij} - \eta_{ij}$. It is given by
\begin{equation}
h_{ij}(t) = H_{ij} e^{(i(\Omega (\bar{n} \cdot \bar{q} - t) + \Phi_0))}
\end{equation}
for phase offset (GW phase at Earth when $t=0$) $\Phi_0$. The amplitude tensor  $H_{ij}$ is
\begin{eqnarray}
	H_{ij} = h_+ e_{ij}^+(\bar{n},\psi) + h_{\times} e_{ij}^{\times}(\bar{n},\psi)
\end{eqnarray}
where $h_{+,\times}$ are the constant amplitudes of the gravitational plane wave, and $e_{ij}^{+, \times}(\bar{n}, \psi)$ are the polarisation tensors which are uniquely defined by the principal axes of the GW. $\psi$ is the polarisation angle of the GW. \newline

@@

\noindent Lets review and categorise all the parameters:
\begin{equation}
	\bar{\theta} =  \bar{\theta}_{\rm PSR} + \bar{\theta}_{\rm GW} + \bar{\theta}_{\rm noise}
\end{equation}
\begin{equation}
	\bar{\theta}_{\rm PSR} = [\gamma, f_{\rm EM}(0), \dot{f}_{\rm EM}(0),d]
\end{equation}
\begin{equation}
	\bar{\theta}_{\rm GW} = [h_{+}, h_{\times}, \delta, \alpha, \psi, \Omega, \Phi_0]
\end{equation}
\begin{equation}
	\bar{\theta}_{\rm noise} = [\sigma, \Sigma]
\end{equation}
We can also express the measurement equation generally as,
\begin{equation} 
	g(\bar{\theta},t) = 1 -A \cos(\Omega(n\cdot q - t) + \Phi_0)
\end{equation}
where 
\begin{equation}
	A =  \frac{1}{2} \frac{H_{ij} q^i q^j}{1 + \bar{n} \cdot \bar{q}}[1 - e^{i \Omega(1+\bar{n} \cdot \bar{q}) d}]
\end{equation}

>>>>

\section{Model}
The intrinsic pulsar frequency $f$ evolves as (See Vargas and Melatos)
\begin{equation}
	\frac{df}{dt} = -\gamma	 [f - f_{\rm EM} (t)] + \dot{f}_{\rm EM} + \xi(t)
	\label{eq:frequency_evolution}
\end{equation}
where $f_{\rm EM}$ is the solution of the electromagnetic spindown equation, $\gamma$ a proporionality constant, and $\xi(t)$ a white noise process that satisfies,
\begin{equation}
	\langle \xi(t) \xi(t') \rangle = \sigma^2 \delta(t - t')
\end{equation}
Over the timescales that we are interested in, we can express the EM spindown simply as
\begin{equation}
	f_{\rm EM} (t) = f_{\rm EM}(0) + \dot{f}_{\rm EM} (0)t
\end{equation}  
Completely, the frequency evolution is then
\begin{equation}
	\frac{df}{dt} = -\gamma	 [f - f_{\rm EM}(0) - \dot{f}_{\rm EM} (0)t] + \dot{f}_{\rm EM} (0) + \xi(t)
	\label{eq:frequency_evolution_expanded}
\end{equation}
This intrinsic frequency can be related to a measured frequency as 
\begin{equation}
	f_M = f g(\bar{\theta},t) + N_M
\end{equation}
where $g(\theta,t)$ ("measurement function")is a function of some parameters, $\bar{\theta}$ and time $t$, whilst $N_M$ is a Gaussian measurement noise that satisfies 
\begin{equation}
	\langle N_M(t) N_M(t') \rangle = \Sigma^2 \delta(t - t')
\end{equation}
Explicitly, it can be shown that the measurement function is,
\begin{equation} \label{eq:final}
	g(\bar{\theta},t) = 1 - \frac{1}{2} \frac{h_{ij}(t) q(t)^i q(t)^j}{1 + \bar{n} \cdot \bar{q}(t)}[1 - e^{i \Omega(1+\bar{n} \cdot \bar{q}(t)) d}]
\end{equation}
There are a few terms to unpack and define in this equation. $q(t)$ is the vector that connects the Earth and the pulsar, $n$ the vector that connects the Earth and the GW source, $\Omega$ the (constant) angular frequency of the GW and $d$ the Earth-pulsar distance. $h_{ij}(t)$ is the metric perturbation due to the GW $= g_{ij} - \eta_{ij}$. It is given by the familiar plane-wave relation:	
\begin{equation}
	h_{ij}(t) = H_{ij} e^{(i(\bar{k} \cdot \bar{x} - \Omega t + \Phi_0))}
\end{equation}
where $\bar{k}$ is the GW 3-wavevector $=\Omega \bar{n}$ and $x = -\bar{q} t$ (photon propagates from pulsar to Earth), and for phase offset (GW phase at Earth when $t=0$) $\Phi_0$.  We can also write this as,
\begin{equation}
h_{ij}(t) = H_{ij} e^{(-i\Omega t (1 + \bar{n} \cdot \bar{q} ) + \Phi_0))}
\end{equation}
The amplitude tensor  $H_{ij}$ is
\begin{eqnarray}
	H_{ij} = h_+ e_{ij}^+(\bar{n},\psi) + h_{\times} e_{ij}^{\times}(\bar{n},\psi)
\end{eqnarray}
where $h_{+,\times}$ are the constant amplitudes of the gravitational plane wave, and $e_{ij}^{+, \times}(\bar{n}, \psi)$ are the polarisation tensors which are uniquely defined by the principal axes of the GW. $\psi$ is the polarisation angle of the GW. \newline

@@

\noindent Lets review and categorise all the parameters:
\begin{equation}
	\bar{\theta} =  \bar{\theta}_{\rm PSR} \cup \bar{\theta}_{\rm GW} \cup \bar{\theta}_{\rm noise}
\end{equation}
\begin{equation}
	\bar{\theta}_{\rm PSR} = [\gamma, f_{\rm EM}(0), \dot{f}_{\rm EM}(0),d]
\end{equation}
\begin{equation}
	\bar{\theta}_{\rm GW} = [h_{+}, h_{\times}, \delta, \alpha, \psi, \Omega, \Phi_0]
\end{equation}
\begin{equation}
	\bar{\theta}_{\rm noise} = [\sigma, \Sigma]
\end{equation}
We can also express the measurement equation generally as,
\begin{equation} 
	g(\bar{\theta},t) = 1 -A \cos(-\Omega t (1 + n\cdot q ) + \Phi_0)
\end{equation}
where 
\begin{equation}
	A =  \frac{1}{2} \frac{H_{ij} q^i q^j}{1 + \bar{n} \cdot \bar{q}}[1 - e^{i \Omega(1+\bar{n} \cdot \bar{q}) d}]
\end{equation}

\cite{Vargas}

\section{PTA pulsars}
